---
title: "Learning Social Media Analytics"
subtitle: "Lecture 1 Course introduction"
author: "Luka Sikic PhD"
institute: "Fakultet hrvatskih studija"
date: "2016/12/12 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```


```{r libs, include=TRUE, echo=FALSE,message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(here)
library(kableExtra)
library(DT)

```


# STRUCTURE
<br>
<br>
<br>
<br>
- General course information
<br>
<br>
- Objectives of the course
<br>
<br>
- Course Literature
<br>
<br>
- Data and resources
<br>
<br>

---
layout: true
# GENERAL COURSE INFORMATION
---
<br>
#### What is this all about?
<br>
<br>
<br>
- understand social media landscape and trends
<br>
<br>
- focus is on methods and techniques for social media analysis
<br>
<br>
- state of the art tools for social media analysis 
<br>
<br>
- independently interpret, analyse or/and develop social media strategy

---
<br>
#### How do we do that?
<br>
<br>
<br>
- R and Python will be be used for the purposes of demonstrating methods
<br>
<br>
- prior knowledge od the programming language syntax is not necessery
<br>
<br>
- learning some of the basic syntax is warmly recommended
<br>
<br>
- we can dedicate one lecture to syntax

---
<br>
#### In the background!
<br>
<br>
- *data science* techniques in focus 
<br>
<br>
- working with databases (*SQL*,*Google Big Querry*)
<br>
<br>
- web data (*web scraping*, *API*)
<br>
<br>
- data manipulation and cleaning (*tidyverse*, *pandas*, *data.table*)
<br>
<br>
- IT ollaboration and code sharing (*Git*, *GitHub*) and open source reporting (*RMarkdown*)
<br>
<br>
- *computational social science* paradigm 

---
<br>
#### Syllabus
<br>
<br>
<br>
- unofficial [syllab](https://raw.githack.com/lusiki/Learning-Social-Media-Analytics/main/syllab/syllab.html) (please follow up for updates); also as [.pdf](https://github.com/lusiki/Learning-Social-Media-Analytics/blob/main/syllab/syllab.pdf).
<br>
<br>
- offiical syllab on [English](https://github.com/lusiki/Learning-Social-Media-Analytics/blob/main/syllab/Learning%20social%20media%20analytics.docx) and [Croatian](https://github.com/lusiki/Learning-Social-Media-Analytics/blob/main/syllab/Analitika%20dru%C5%A1tvenih%20medija.docx)

---
layout: true
# OBJECTIVES OF THE COURSE
---
<br>
<br>
<br>

- Catching up with the current state and trends in the social media space 
<br>
<br>
- IT infrastructure (Big Data) of social media platforms  
<br>
<br>
- Relevant topics in social media research and business 
<br>
<br>
- Tools, methodology and results in social media analytics
<br>
<br>
- Independent social media research

---
layout: true
# COURSE LITERATURE
---
<br>
#### Books
<br>
<br>
- [Mining the Social Web](https://www.webpages.uidaho.edu/~stevel/504/mining-the-social-web-2nd-edition.pdf) 
<br>
<br>
- [Learning Social Media Analytics with R](https://www.packtpub.com/product/learning-social-media-analytics-with-r/9781787127524)
<br>
<br>
- [Mastering Social Media Mining with Python](https://www.packtpub.com/product/mastering-social-media-mining-with-python/9781783552016) 
<br>
<br>
- [Social Media Data Mining and Analytics](https://www.wiley.com/en-us/Social+Media+Data+Mining+and+Analytics-p-9781118824856) 
<br>
<br>
- [Social Media Analytics Strategy](https://www.apress.com/us/book/9781484231012)
---
<br>
#### Academic literature
<br>
<br>
- [International Conference on Web and Social Media](https://ojs.aaai.org/index.php/ICWSM/issue/archive)
<br>
<br>
- [Social Media + Society](https://journals.sagepub.com/home/sms)
<br>
<br>
- [Information, Communication & Society](https://www.tandfonline.com/journals/rics20)
<br>
<br>
- Optionally you can pick another journal to follow, check the [list](https://lusiki.github.io/Learning-Social-Media-Analytics/resources.html) 
---
layout: true
# DATA AND RESOURCES
---
<br>
#### Data
<br>
<br>
- There are numerous social media data sources
<br>
<br>
- We will be working with Croatian data so please thoughly check [Mediatoolkit](https://www.mediatoolkit.com/) (alternatives are [presscut](https://www1.presscut.hr/en/), [pressclipping](https://www.pressclipping.hr/))
<br>
<br>
- Data samples are provided in the repo
<br>
<br>
- These data sources are equally inconvenient (expensive and partitial)
<br>
<br>
- Serious solution requires your own database complilation 


---
<br>
#### Resources 
<br>
<br>
- There are countless social media resources
<br>
<br>
- Please check my [curated list](https://lusiki.github.io/Learning-Social-Media-Analytics/resources.html)
<br>
<br>
- Consider adding to it (maybe as a student project)

---
class: inverse, middle
layout:false
# Thank you for your attention!
