---
title: "Learning Social Media Analytics"
subtitle: "Current state and latest trends"
author: "Luka Sikic, PhD"
institute: "Faculty of Croatian Studies"
date: "(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```



```{r libs, include=TRUE, echo=FALSE,message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(here)
library(kableExtra)
library(DT)
library(purrr)
library(data.table)
library(RMySQL)
```


# CLASS OUTLINE
<br>
<br>
- DATA OVERVIEW (full media space)
<br>
<br>
- FACEBOOK
<br>
<br>
- WEB
<br>
<br>
- TWITTER
<br>
<br>
- YOUTUBE
<br>
<br>
- FORUM
<br>
<br>
- INSTAGRAM
<br>
<br>

---
layout: true
# DATA OVERVIEW (full media space)
---

- check the raw [data](https://github.com/lusiki/Learning-Social-Media-Analytics/blob/main/data/Mediatoolkit/sample.xlsx)
<br>
- sourced from [Mediatoolkit](https://www.mediatoolkit.com/)



```{r, echo=F}
# read data in (one day worth of data)
one_day_sample <- read_excel(here::here("data/Mediatoolkit/sample.xlsx"))
```

- one day worth of data (2022-01-01)

```{r, echo=F}
dim(one_day_sample)
```

- random data snapshot

```{r, echo=F}
glimpse(one_day_sample[sample(nrow(one_day_sample)),1:10])
```
---

- avaliable data characteristics

<br>

```{r, echo=F}
# other variables
names(one_day_sample)
```

---
<br>

- rows and columns 2021


```{r, eval = F, warning=FALSE, message=FALSE,  echo=F, cache = TRUE}
conn <- dbConnect(RMySQL::MySQL(), dbname = "determ_all", host = "127.0.0.1",
                  user = "Lux", password = "Theanswer0207", local_infile = TRUE)

query <- "SELECT * FROM media_space_2023"
data <- dbGetQuery(conn, query)


#fwrite(data, "C:/Users/lukas/Desktop/dta23.xlsx")

```
```{r}
data_read <- fread("C:/Users/lukas/Desktop/dta23.xlsx")
```


- time span

```{r, echo=F}
range(data_read$DATE)
```

- social media avaliable

```{r, echo=F}
unique(data_read$SOURCE_TYPE)
```

- how many authors

```{r, echo=F}
length(unique(data_read$AUTHOR))
```

- how many unique titles

```{r, echo=F}
# unique titles
length(unique(data_read$TITLE))
```

---

- how many unique sources

```{r, echo=F}
# unique sources
length(unique(data_read$FROM))
```

- how many unique urls

```{r, echo=F}
# unique urls
length(unique(data_read$URL))

```

---

- how much activity per media

<br>
<br>
```{r, echo=F, warning=FALSE, message=FALSE}
# GENERAL OVERVIEW 
data_read %>%
  group_by(SOURCE_TYPE) %>%
  count() %>%
  arrange(desc(n)) %>%
  mutate_if(is.numeric, format, big.mark = ".") %>%
  kable(., "html") %>%
  kable_styling("striped", full_width = F) 
  
```


---

- check some text (random choice)

```{r, echo=F, warning=FALSE, message=FALSE}
# SAMPLE SOME TXT
data_read %>%
  select(SOURCE_TYPE, MENTION_SNIPPET, URL) %>%
  unique() %>%
  sample_n(50) %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")

```

---

- check some titles (random choice)
<br> 
<br>
```{r, echo=F, warning=FALSE, message=FALSE}
# SAMPLE SOME TITLES
data_read %>%
  select(SOURCE_TYPE, TITLE, URL) %>%
  sample_n(50)  %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
```






---
layout: true
# FACEBOOK
---
<br>


```{r, echo=F, warning=FALSE, message=FALSE}
# choose variables to investigate
data_read %>%
  filter(SOURCE_TYPE == "facebook") %>%
  select(AUTHOR,
         REACH,
         FOLLOWERS_COUNT,
         LIKE_COUNT,
         COMMENT_COUNT,
         TOTAL_REACTIONS_COUNT,
         INFLUENCE_SCORE,
         TITLE,
#         MENTION_SNIPPET,
         DATE, 
         TIME) -> fb
```


- most important authors (per followers)
<br>

```{r echo=F}
fb %>%
  group_by(AUTHOR) %>% 
  summarize(FOLLOWERS_COUNT = max(FOLLOWERS_COUNT)) %>% 
  ungroup() %>%
  arrange(desc(FOLLOWERS_COUNT)) %>%
  head(50) %>%
  mutate_if(is.numeric, format, big.mark = ",") %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
```

---

- most important authors (per followers)
<br>

```{r, echo=F, warning=FALSE, message=FALSE}
# most active authors
fb %>% 
  group_by(AUTHOR) %>%
  summarise( N = n()) %>%
  ungroup() %>%
  arrange(desc(N)) %>%
  mutate(N = format(N, big.mark = ",")) %>%
  head(1000)  %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")

```


---

- most important authors (per likes)
<br>
<br>
```{r, echo=F, warning=FALSE, message=FALSE}
# authors with impact
fb %>% 
  group_by(AUTHOR) %>%
  summarise(LIKES = sum(LIKE_COUNT),
            COMMENT = sum(COMMENT_COUNT),
            REACH = sum(REACH),
            REACTIONS = sum(TOTAL_REACTIONS_COUNT)) %>%
  arrange(desc(LIKES)) %>%
  mutate_if(is.numeric, format, big.mark = ",") %>%
  head(100)  %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
```


---

- summary of  important activity  
<br>
```{r, echo=F, warning=FALSE, message=FALSE}

fb %>% 
  group_by(TITLE) %>%
  summarise(LIKES = sum(LIKE_COUNT),
            COMMENT = sum(COMMENT_COUNT),
            REACH = sum(REACH),
            REACTIONS = sum(TOTAL_REACTIONS_COUNT)) %>%
  select(TITLE, LIKES, COMMENT, REACH,REACTIONS ) %>%
  arrange(desc(LIKES)) %>%
  head(50) %>%
  mutate_if(is.numeric, format, big.mark = ",") %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
```

---

- activity of important authors  

<br>

```{r, echo=F, warning=FALSE, message=FALSE}
# important activity and author
fb %>% 
  group_by(AUTHOR) %>%
  mutate(LIKES = sum(LIKE_COUNT),
         COMMENT = sum(COMMENT_COUNT),
         REACH = sum(REACH),
         REACTIONS = sum(TOTAL_REACTIONS_COUNT)) %>%
  select(AUTHOR, TITLE, LIKES, COMMENT, REACH,REACTIONS ) %>%
  arrange(desc(LIKES)) %>%
   slice(50) %>%
  mutate_if(is.numeric, format, big.mark = ",") %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
```


---
layout: true
# WEB
---
<br>

```{r, echo=F, warning=FALSE, message=FALSE}
# choose variables to investigate
data_read %>%
  filter(SOURCE_TYPE == "web") %>%
  select(FROM_SITE,
         AUTHOR,
         REACH,
         VIRALITY,
         LIKE_COUNT,
         COMMENT_COUNT,
         INFLUENCE_SCORE,
         TITLE,
#       MENTION_SNIPPET,
         DATE, 
         TIME) -> web
```

- most active authors
  
```{r, echo=F, warning=FALSE, message=FALSE}
# most active authors
web %>% 
  group_by(FROM_SITE) %>%
  summarise(N = n()) %>%
  arrange(desc(N))  %>%
  head(1000) %>%
  mutate_if(is.numeric, format, big.mark = ".") %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
```

---

- summary of important sources (by likes)
  
```{r, echo=F, warning=FALSE, message=FALSE}
web %>% 
  group_by(FROM_SITE) %>%
  summarise(LIKES = sum(LIKE_COUNT),
            COMMENT = sum(COMMENT_COUNT),
            REACH = sum(REACH),
            VIRALITY = sum(VIRALITY)) %>%
  arrange(desc(LIKES)) %>%
  head(1000) %>%
  mutate_if(is.numeric, format, big.mark = ",")  %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
```


---
layout: true
# TWITTER
---
  
```{r, echo=F, warning=FALSE, message=FALSE,}
# choose variables to investigate
data_read %>%
  filter(SOURCE_TYPE == "twitter") %>%
  select(AUTHOR,
         REACH,
         FOLLOWERS_COUNT,
         FAVORITE_COUNT,
         RETWEET_COUNT,
         TWEET_TYPE,
         TWEET_SOURCE_NAME,
         TITLE,
         #      MENTION_SNIPPET,
         DATE, 
         TIME) -> twitter
```

- most important authors (by followers)
  
```{r, echo=F, warning=FALSE, message=FALSE,}
# most important authors
twitter %>% 
  group_by(AUTHOR) %>%
  summarise(FOLLOWERS = max(FOLLOWERS_COUNT)) %>%
  arrange(desc(FOLLOWERS))  %>%
  head(50) %>%
  mutate_if(is.numeric, format, big.mark = ",")  %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
```

---

- most active authors

```{r, echo=F, warning=FALSE, message=FALSE,}
# most active authors
twitter %>% 
  group_by(AUTHOR) %>%
  summarise(N = n()) %>%
  arrange(desc(N))  %>%
  head(50) %>%
  mutate_if(is.numeric, format, big.mark = ",")  %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
```

---

- summary of most active authors

```{r, echo=F, warning=FALSE, message=FALSE,}
# most active authors
twitter %>% 
  group_by(AUTHOR) %>%
  summarise(FAVORITE = sum(FAVORITE_COUNT),
            RETWEET = sum(RETWEET_COUNT)) %>%
  arrange(desc(RETWEET))  %>%
  head(50) %>%
  mutate_if(is.numeric, format, big.mark = ",") %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
```

---
 
- originality of conntent
<br>
<br>
```{r, echo=F, warning=FALSE, message=FALSE,}
# originality
twitter %>% 
  group_by(TWEET_TYPE) %>% # interesting to check per author
  summarise(N = n()) %>%
  arrange(desc(N)) %>%
  mutate_if(is.numeric, format, big.mark = ",") %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
```

---

- technology used
<br>
<br>
```{r, echo=F, warning=FALSE, message=FALSE,}
# technology
twitter %>% 
  group_by(TWEET_SOURCE_NAME) %>%
  summarise(N = n()) %>%
  arrange(desc(N)) %>%
  mutate_if(is.numeric, format, big.mark = ",")  %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
```

---
layout: true
# YOUTUBE
---
<br>
<br>
```{r, echo=F, warning=FALSE, message=FALSE,}
# choose variables to investigate
data_read %>%
  filter(SOURCE_TYPE == "youtube") %>%
  select(AUTHOR,
         REACH,
         LIKE_COUNT,
         VIEW_COUNT,
         COMMENT_COUNT,
         TITLE,
#        MENTION_SNIPPET,
         URL,
         DATE, 
         TIME) -> youtube
```

- most active youtubers
<br>
<br>
```{r, echo=F, warning=FALSE, message=FALSE,}
# most active authors
youtube %>%
  group_by(AUTHOR) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(1000) %>%
  mutate_if(is.numeric, format, big.mark = ",") %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")

```

---

- summary of activity (per view)
<br>
<br>
```{r, echo=F, warning=FALSE, message=FALSE,}
# activity per author
youtube %>%
  group_by(AUTHOR) %>%
  summarise(REACH = sum(REACH),
            LIKE = sum(LIKE_COUNT),
            VIEW = sum(VIEW_COUNT),
            COMMENTS= sum(COMMENT_COUNT)) %>%
  arrange(desc(VIEW)) %>%
  head(1000) %>%
  mutate_if(is.numeric, format, big.mark = ",") %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
``` 

---

- conntent summary (most views)
<br>
<br>
```{r, echo=F, warning=FALSE, message=FALSE,}
# breaks down
youtube %>%
  group_by(AUTHOR, URL,REACH,LIKE_COUNT, VIEW_COUNT, COMMENT_COUNT) %>%
  arrange(desc(VIEW_COUNT)) %>%
  head(1000) %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")

```   

---
layout: true
# FORUM
---
<br>
<br>

```{r, echo=F, warning=FALSE, message=FALSE,}
# choose variables to investigate
data_read %>%
  filter(SOURCE_TYPE == "forum") %>%
  select(FROM_SITE,
         TITLE,
         #      MENTION_SNIPPET,
         DATE, 
         TIME) -> forum
``` 

- not a very good data coverage!?
<br>
<br>
```{r, echo=F, warning=FALSE, message=FALSE,}
# not a very good coverage
forum %>%
  group_by(FROM_SITE) %>%
  count() %>%
  arrange(desc(n)) %>%
  mutate_if(is.numeric, format, big.mark = ".") %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
``` 

---
layout: true
# INSTAGRAM
---
<br>

```{r, echo=F, warning=FALSE, message=FALSE,}
# choose variables to investigate
data_read %>%
  filter(SOURCE_TYPE == "instagram") %>%
  select(FROM_SITE,
         TITLE,
         AUTHOR,
         #      MENTION_SNIPPET,
         REACH,
         LIKE_COUNT,
         COMMENT_COUNT,
         SHARE_COUNT,
         VIEW_COUNT,
         DATE, 
         TIME) -> instagram
``` 

- influencer statistics
<br> 
<br>
```{r, echo=F, warning=FALSE, message=FALSE,}
# influencer impact
instagram %>%
  group_by(FROM_SITE) %>%
  summarise(REACH = sum(REACH),
            COMMENT = sum(COMMENT_COUNT),
            LIKE = sum(LIKE_COUNT),
            VIEW = sum(VIEW_COUNT)) %>%
  mutate_if(is.numeric, format, big.mark = ",") %>%
  arrange(desc(LIKE)) %>%
  head(100)  %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
```   

---

- influencer activity
<br>
<br>
```{r, echo=F, warning=FALSE, message=FALSE,}
# influencer activity
instagram %>%
  group_by(FROM_SITE) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(100) %>%
  mutate_if(is.numeric, format, big.mark = ".") %>%
  kable("html") %>% # Specify HTML to use kableExtra features effectively
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), # Make table responsive
                full_width = F, 
                fixed_thead = T) %>% # Keep the header visible while scrolling
  scroll_box(width = "100%", height = "400px")
```    



---
layout: true
# RESOURCES
---
<br>
<br>
<br>
- The data sample in the GiHub repo
<br>
<br>
- Check other clipping agencies
<br>
<br>
- Other data vendors
<br>
<br>
- DATAREPORTAL

---
class: inverse, middle
layout:false
# Thank you for your attention!































