---
title: "Learning Social Media Analytics"
# subtitle: "<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>"
subtitle: "Lecture 3: Data Science Prerequisites"
author: "Luka Sikic, PhD"
date: "Faculty of Croatian Studies | [LSMA](https://lusiki.github.io/Learning-Social-Media-Analytics/)" #"`r format(Sys.time(), '%d %B %Y')`"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts] 
    lib_dir: libs
    nature:
      highlightStyle: googlecode
      highlightLines: true
      countIncrementalSlides: false
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
```


```{r libs, include=TRUE, echo=FALSE,message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(here)
library(kableExtra)
library(DT)
```


# STRUCTURE
<br>
<br>
- FETCHING DATA
<br>
<br>
- STORING DATA
<br>
<br>
- MANIPULATING DATA
<br>
<br>
- ANALYTICS
<br>
<br>
-  REPORTING


---
layout: true
# FETCHING DATA
---



```{r, echo=T}


```


---
<br>
<br>

```{r, echo=T}

```

---
<br>
<br>
```{r, echo=T}

```

---

```{r, echo=T}

```

---

```{r, echo=T}

```

---

```{r, echo=T}

```

---

```{r, echo=T}

```

---
layout: true
# STORING DATA
---
<br>
<br>
```{r , echo=T}

```

---

```{r, echo=T}

```

---

```{r, echo=T}

```

---

```{r, echo=T}

```


```{r, echo=F, eval=F, include=F}

```

---
layout: true
# MANIPULATING DATA
---
<br>
<br>

```{r , echo=T}

```

---

```{r, echo=T}

```
  
---

```{r, echo=T}

```

---
layout: true
# ANALYTICS
---

```{r, echo=T}

```
  
---

```{r, echo=T}


```

---

```{r, echo=T}

```

---

```{r, echo=T}

```
  
---

```{r, echo=T}

```

---

```{r, echo=T}

```
  
---
layout: true
# REPORTING
---
<br>
<br>
```{r, echo=T}

```
  
---

```{r, echo=T}


```

---

```{r, echo=T}

``` 
  
  
---
<br>
#### Resources
<br>
<br>
- Check the data sample in GHub repo
<br>
<br>
- Check other clipping agencies
<br>
<br>
- Check other data vendors

---
class: inverse, middle
layout:false
# Thank you for your attention!